---
title: "forestChange"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r}
library(sp)
library(forestChange)
```

## Download Hansen data

Download tree cover in 2000 and the year of forest loss for French Guiana:

```{r}
GUF_poly <- getGADM(level = 0, country = "GUF")
GUF_data <- FCPolygon(pol = GUF_poly)
```

Plot the data:

```{r}
plot(GUF_data[[1]])
plot(GUF_data[[2]])
```

## Calculate EBV metrics

The package can calculate forest extents, fractal-dimension indices, and other values in `PatchStat`. It filters out regions with forest cover values lower than the argument `perc` and then removes forest cover values from the year 2000 in regions where forest was lost during the years 2000 to `year`.

```{r}
forest_ext <- EBVmetric(GUF_data, met = "forest.ext", year = 16, perc = 60)
# Warning message:
# In EBVmetric(GUF_data, met = "forest.ext", year = 16, perc = 60) :
#    Year 16 has no GFC records.
frac_dim <- EBVmetric(GUF_data, met = "frac.dim.index", year = 3, perc = 60)
```

Looking at the NA values in the loss year raster, we can see that 2014 is the latest year we can get EBV metrics for:

```{r}
lss <- GUF_data[[names(GUF_data)[grepl('loss', names(GUF_data))]]]
myr <- max(na.omit(values(lss)))
myr
```


