---
title: "forestChange"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r}
library(sp)
library(forestChange)
```

## Download Hansen data

Download tree cover in 2000 and the year of forest loss for French Guiana:

```{r}
# level = 0 gets entire country
# country name is ISO code speecifying the country
GUF_poly <- getGADM(level = 0, country = "GUF")
# can get layers treecover2000, loss, gain, lossyear, datamask, first, and last
GUF_data <- FCPolygon(pol = GUF_poly)
```

Plot the data:

```{r}
plot(GUF_data[[1]], main = "Tree Cover 2000")
plot(GUF_data[[2]], main = "Year of Forest Loss")
```

## Calculate EBV metrics

The package can calculate forest extents, fractal-dimension indices, and other values in `PatchStat`. It filters out regions with forest cover values lower than the argument `perc` and then removes forest cover values from the year 2000 in regions where forest was lost during the years 2000 to `year`.

```{r}
forest_ext <- EBVmetric(GUF_data, met = "forest.ext", year = 13, perc = 60)
frac_dim <- EBVmetric(GUF_data, met = "frac.dim.index", year = 3, perc = 60)
```

Looking at the NA values in the loss year raster, we can see that 2014 is the latest year we can get EBV metrics for:

```{r}
lss <- GUF_data[[names(GUF_data)[grepl('loss', names(GUF_data))]]]
myr <- max(na.omit(values(lss)))
myr
```

## Combining multiple countries

Download the data for Suriname:

```{r}
# level = 0 gets entire country
# country name is ISO code speecifying the country
SUR_poly <- getGADM(level = 0, country = "SUR")
SUR_data <- FCPolygon(pol = SUR_poly)
```

Plot the data:

```{r}
plot(SUR_data[[1]], main = "Tree Cover 2000")
plot(SUR_data[[2]], main = "Year of Forest Loss")
```

Use `merge` to combine the French Guiana and Suriname layers:

```{r}
GUF_SUR_data <- merge(GUF_data, SUR_data)
names(GUF_SUR_data) <- names(GUF_data)
plot(GUF_SUR_data[[1]], main = "Tree Cover 2000")
plot(GUF_SUR_data[[2]], main = "Year of Forest Loss")
```

## Calculate EBV metrics

We can calculate EBV metrics for the combined RasterStack as well.

```{r}
forest_ext <- EBVmetric(GUF_SUR_data, met = "forest.ext", year = 13, perc = 60)
frac_dim <- EBVmetric(GUF_SUR_data, met = "frac.dim.index", year = 3, perc = 60)
```

Looking at the NA values in the loss year raster, we can see that 2014 is the latest year we can get EBV metrics for:

```{r}
lss <- GUF_data[[names(GUF_data)[grepl('loss', names(GUF_data))]]]
myr <- max(na.omit(values(lss)))
myr
```