---
title: "GBIF occurrence points"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Including GBIF points

Since the occurrence data I used to create the Maxent models is all historical data from different papers, I currently have no occurrence points recent enough to temporally match with remotely sensed forest cover data. To address this issue, I have been looking into using GBIF occurrence data to find a forest cover threshold. The quality of the GBIF data is fairly poor though, so I need to develop some kind of standard for whether a given GBIF point is close enough to my known occurrence points for that species to be included.

```{r, message=FALSE}
# load packages
library(spocc)
library(ggmap)
library(rgeos)
library(sp)
api_key = "AIzaSyBK7lLbqoqnYFdzf-idYYposb-1gwyRAlQ"
register_google(key = api_key)
```

### Literature data

The following is the occurrence data from papers that I have used to create my Maxent models.

```{r, message=FALSE}
# read in literature data
dataDir = '/Users/hellenfellows/OneDrive\ -\ AMNH/Wallace/Occurrence_Data/'

# Species occurrence coordinates
variegatus_lit <- read.csv(paste0(dataDir,'Bradypus_variegatus_litdata.csv'))
tridactylus_lit <- read.csv(paste0(dataDir,'Bradypus_tridactylus_litdata.csv'))
torquatus_lit <- read.csv(paste0(dataDir,'Bradypus_torquatus_litdata.csv'))

SA_bbox <- make_bbox(lon = c(-97, -25), lat = c(-25,20), f = 0.1)
SA_map <- get_map(location = SA_bbox, source = "google", maptype = "satellite")
ggmap(SA_map) +
  ggtitle("Literature Occurrence Points") +
  geom_point(data = variegatus_lit, aes(x = longitude, y = latitude, color = "variegatus"), cex = 0.8) +
  geom_point(data = tridactylus_lit, aes(x = longitude, y = latitude, color = "tridactylus"), cex = 0.8) +
  geom_point(data = torquatus_lit, aes(x = longitude, y = latitude, color = "torquatus"), cex = 0.8) +
  scale_color_manual(name = "Legend", values = c(variegatus = "darkorange1", tridactylus = "springgreen", torquatus = "deepskyblue"))
```

### GBIF data

I loaded GBIF data for all three species and removed records with no coordinates as well as records at the coordinates (0, 0).

```{r}
variegatus_gbif <- occ(query = "Bradypus variegatus", from = 'gbif', limit = 2000)
tridactylus_gbif <- occ(query = "Bradypus tridactylus", from = 'gbif', limit = 2000)
torquatus_gbif <- occ(query = "Bradypus torquatus", from = 'gbif', limit = 2000)

variegatus_gbif <- occ2df(variegatus_gbif)
tridactylus_gbif <- occ2df(tridactylus_gbif)
torquatus_gbif <- occ2df(torquatus_gbif)

# remove NAs
variegatus_gbif <- variegatus_gbif[!is.na(variegatus_gbif$longitude), ]
tridactylus_gbif <- tridactylus_gbif[!is.na(tridactylus_gbif$longitude), ]
torquatus_gbif <- torquatus_gbif[!is.na(torquatus_gbif$longitude), ]

# remove points at 0, 0
variegatus_gbif <- variegatus_gbif[variegatus_gbif$longitude != 0, ]
tridactylus_gbif <- tridactylus_gbif[tridactylus_gbif$longitude != 0, ]
torquatus_gbif <- torquatus_gbif[torquatus_gbif$longitude != 0, ]

ggmap(SA_map) +
  ggtitle("GBIF Occurrence Points") +
  geom_point(data = variegatus_gbif, aes(x = longitude, y = latitude, color = "variegatus"), cex = 0.8) +
  geom_point(data = tridactylus_gbif, aes(x = longitude, y = latitude, color = "tridactylus"), cex = 0.8) +
  geom_point(data = torquatus_gbif, aes(x = longitude, y = latitude, color = "torquatus"), cex = 0.8) +
  scale_color_manual(name = "Legend", values = c(variegatus = "darkorange1", tridactylus = "springgreen", torquatus = "deepskyblue"))
```

Clearly, certain species, particularly *B. tridactylus*, are frequently misidentified, with occurrence records well outside of the species' range.

Here, we can see maps for each species with the data from literature and the GBIF points in red:

```{r}
ggmap(SA_map) +
  ggtitle("Bradypus variegatus") +
  geom_point(data = variegatus_lit, aes(x = longitude, y = latitude, color = "literature"), cex = 0.8) +
  geom_point(data = variegatus_gbif, aes(x = longitude, y = latitude, color = "GBIF"), cex = 0.8) +
  scale_color_manual(name = "Legend", values = c(literature = "darkorange1", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus tridactylus") +
  geom_point(data = tridactylus_lit, aes(x = longitude, y = latitude, color = "literature"), cex = 0.8) +
  geom_point(data = tridactylus_gbif, aes(x = longitude, y = latitude, color = "GBIF"), cex = 0.8) +
  scale_color_manual(name = "Legend", values = c(literature = "springgreen", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus torquatus") +
  geom_point(data = torquatus_lit, aes(x = longitude, y = latitude, color = "literature"), cex = 0.8) +
  geom_point(data = torquatus_gbif, aes(x = longitude, y = latitude, color = "GBIF"), cex = 0.8) +
  scale_color_manual(name = "Legend", values = c(literature = "deepskyblue", GBIF = "red"))
```


## Minimum convex polygon

The first method I tried to identify GBIF points "close enough" to known occurrence points was a minimum convex polygon (MCP) for each species.

```{r}
# function to make a minimum convex polygon as SpatialPolygons object
mcp <- function(xy) {
  xy <- as.data.frame(sp::coordinates(xy))
  coords.t <- chull(xy[, 1], xy[, 2])
  xy.bord <- xy[coords.t, ]
  xy.bord <- rbind(xy.bord[nrow(xy.bord), ], xy.bord)
  return(sp::SpatialPolygons(list(sp::Polygons(list(sp::Polygon(as.matrix(xy.bord))), 1))))
}

# create MCP for literature data from all three species
var_mcp <- mcp(variegatus_lit[,2:3])
tri_mcp <- mcp(tridactylus_lit[,2:3])
tor_mcp <- mcp(torquatus_lit[,2:3])
```

Now we can visualize the MCP for each species with the occurrence points from literature and GBIF:

```{r}
ggmap(SA_map) +
  ggtitle("Bradypus variegatus") +
  geom_point(data = variegatus_lit, aes(x = longitude, y = latitude, color = "literature"), cex = 0.8) +
  geom_point(data = variegatus_gbif, aes(x = longitude, y = latitude, color = "GBIF"), cex = 0.8) +
  geom_polygon(data = fortify(var_mcp), aes(x = long, y = lat), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "darkorange1", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus tridactylus") +
  geom_point(data = tridactylus_lit, aes(x = longitude, y = latitude, color = "literature"), cex = 0.8) +
  geom_point(data = tridactylus_gbif, aes(x = longitude, y = latitude, color = "GBIF"), cex = 0.8) +
  geom_polygon(data = fortify(tri_mcp), aes(x = long, y = lat), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "springgreen", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus torquatus") +
  geom_point(data = torquatus_lit, aes(x = longitude, y = latitude, color = "literature"), cex = 0.8) +
  geom_point(data = torquatus_gbif, aes(x = longitude, y = latitude, color = "GBIF"), cex = 0.8) +
  geom_polygon(data = fortify(tor_mcp), aes(x = long, y = lat), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "deepskyblue", GBIF = "red"))
```

Overall, the MCP approach includes too many of the GBIF points for *B. variegatus*, particularly the occurrence point that is in *B. tridactylus*'s range. The MCP for *B. tridactylus* looks fairly good, although it might be a little overly conservative by excluding all of the *B. tridactylus* GBIF points along the French Guiana - Brazil border. The MCP for torquatus looks good -- it excludes the GBIF points I'd like to get rid of.

## Buffered points

Next I tried some buffered point regions, starting at 1 degree and moving to 4 degrees (since a 4 degree buffer was what we selected as a background region for Maxent modeling).

### 1 degree buffer

```{r}
# make SpatialPoints objects for buffering
var_sp <- SpatialPoints(variegatus_lit[,2:3])
tri_sp <- SpatialPoints(tridactylus_lit[,2:3])
tor_sp <- SpatialPoints(torquatus_lit[,2:3])

var_buffer <- gBuffer(var_sp, width = 1)
tri_buffer <- gBuffer(tri_sp, width = 1)
tor_buffer <- gBuffer(tor_sp, width = 1)

ggmap(SA_map) +
  ggtitle("Bradypus variegatus") +
  geom_point(data = variegatus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = variegatus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = var_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "darkorange1", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus tridactylus") +
  geom_point(data = tridactylus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = tridactylus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = tri_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "springgreen", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus torquatus") +
  geom_point(data = torquatus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = torquatus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = tor_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "deepskyblue", GBIF = "red"))
```

The 1-degree buffer is definitely much more conservative than the MCP. For *B. torquatus*, the 1-degree buffer still includes the points that I would want to include. I'm fairly comfortable with this conservative estimate for *B. variegatus* too. The 1-degree buffer region for *B. tridactylus*, however, excludes a lot of occurrence points that I would guess are true occurrence points.

## 2 degree buffer

```{r}
var_buffer <- gBuffer(var_sp, width = 2)
tri_buffer <- gBuffer(tri_sp, width = 2)
tor_buffer <- gBuffer(tor_sp, width = 2)

ggmap(SA_map) +
  ggtitle("Bradypus variegatus") +
  geom_point(data = variegatus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = variegatus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = var_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "darkorange1", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus tridactylus") +
  geom_point(data = tridactylus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = tridactylus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = tri_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "springgreen", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus torquatus") +
  geom_point(data = torquatus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = torquatus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = tor_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "deepskyblue", GBIF = "red"))
```

I'm pretty comfortable with the 2-degree buffer region for all three species, although it may still be too conservative for *B. tridactylus*.

### 4 degree buffer

```{r}
var_buffer <- gBuffer(var_sp, width = 4)
tri_buffer <- gBuffer(tri_sp, width = 4)
tor_buffer <- gBuffer(tor_sp, width = 4)

ggmap(SA_map) +
  ggtitle("Bradypus variegatus") +
  geom_point(data = variegatus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = variegatus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = var_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "darkorange1", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus tridactylus") +
  geom_point(data = tridactylus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = tridactylus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = tri_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "springgreen", GBIF = "red"))

ggmap(SA_map) +
  ggtitle("Bradypus torquatus") +
  geom_point(data = torquatus_lit, aes(x = longitude, y = latitude, color = "literature")) +
  geom_point(data = torquatus_gbif, aes(x = longitude, y = latitude, color = "GBIF")) +
  geom_polygon(data = tor_buffer, aes(x = long, y = lat, group = group), color = "black", alpha = 0) +
  scale_color_manual(name = "Legend", values = c(literature = "deepskyblue", GBIF = "red"))
```

The 4-degree buffer definitely includes too many GBIF points for all 3 species.

## Conclusions

Right now, I'm leaning towards using the 1-degree point buffer or possibly the 2-degree buffer as my criterion for GBIF points to be used in calculating forest cover threshold. The MCP and 4-degree buffers include too many of the erroneous GBIF points and should not be used. Alternatively, I could do an alpha-hull instead of the MCP, but since we only need enough recent points to find a forest cover threshold, I don't think we need to be too concerned about including every single correct GBIF occurrence point. In the spirit of being cautious, I'm thinking I move forward with finding forest cover values for the GBIF points within a 1-degree buffer of the literature occurrence points -- let me know if this sounds good or if there is a different way I should try out.

