---
title: "MaskedSVM_SDMs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages and source functions

```{r}
library(raster)
library(maskRangeR)
```


## Load occurrence data and SDMs

I loaded the occurrence data for the three species, with latitude and longitude points coming from literature.

```{r}
variegatus <- read.csv("~/OneDrive - AMNH/Wallace/Occurrence_Data/Bradypus_variegatus_litdata.csv")
head(variegatus)
tridactylus <- read.csv("~/OneDrive - AMNH/Wallace/Occurrence_Data/Bradypus_tridactylus_litdata.csv")
head(tridactylus)
torquatus <- read.csv("~/OneDrive - AMNH/Wallace/Occurrence_Data/Bradypus_torquatus_litdata.csv")
head(torquatus)
```

I used SDMs built from a 4-degree background buffer around points thinned to 40 km. Models were selected based on the lowest AIC and lowest omission rate.

```{r}
#load SDMs
var_sdm <- raster("/Users/hellenfellows/OneDrive\ -\ AMNH/Wallace/Wallace_code/thinned_variegatus_L_1_cloglog.tif")
plot(var_sdm)
tri_sdm <- raster("/Users/hellenfellows/OneDrive\ -\ AMNH/Wallace/Wallace_code/thinned_tridactylus_LQ_2_cloglog.tif")
plot(tri_sdm)
tor_sdm <- raster("/Users/hellenfellows/OneDrive\ -\ AMNH/Wallace/Wallace_code/thinned_torquatus_H_3_cloglog.tif")
plot(tor_sdm)
```

## Spatial SVMs

The following SVMs are built using only spatial information, i.e. the occurrence localities of the two species.

### Unweighted spatial SVMs

In the unweighted SVM, the occurrences of each species are weighed equally.

```{r, message = FALSE}
# create SVM
unweighted_svm_occ <- rangeSVM(variegatus[,2:3], tridactylus[,2:3], torquatus[,2:3], nrep = 100)
```

```{r}
# use SVM to create a raster of predicted regions
sloth_unweighted_svm_occ <- rangeSVM_predict(svm = unweighted_svm_occ, r = var_sdm)
plot(sloth_unweighted_svm_occ)
points(variegatus[,2:3], pch = 20, cex = 0.75, col = "black")
points(tridactylus[,2:3], pch = 20, cex = 0.75, col = "red")
points(torquatus[,2:3], pch = 20, cex = 0.75, col = "blue")
```

The grey areas indicate regions predicted for *B. variegatus*, the yellow areas indicate regions predicted for *B. tridactylus*, and the green areas indicate regions predicted for *B. torquatus*. Black points show the occurrences of *B. variegatus*, red show *B. tridactylus*, and blue show *B. torquatus*.

#### *B. variegatus*

```{r}
# masked SDM predictions for variegatus
var_unweighted_svm_occ <- sloth_unweighted_svm_occ == 1
var_unweighted_svm_occ[var_unweighted_svm_occ == 0] <- NA
var_unweighted_svm_occ_mask <- mask(var_sdm, var_unweighted_svm_occ)
plot(var_unweighted_svm_occ_mask)
```

#### *B. tridactylus*

```{r}
# masked SDM predictions for tridactylus
tri_unweighted_svm_occ <- sloth_unweighted_svm_occ == 2
tri_unweighted_svm_occ[tri_unweighted_svm_occ == 0] <- NA
tri_unweighted_svm_occ_mask <- mask(tri_sdm, tri_unweighted_svm_occ)
plot(tri_unweighted_svm_occ_mask)
```

#### *B. torquatus*

```{r}
# masked SDM predictions for torquatus
tor_unweighted_svm_occ <- sloth_unweighted_svm_occ == 3
tor_unweighted_svm_occ[tor_unweighted_svm_occ == 0] <- NA
tor_unweighted_svm_occ_mask <- mask(tor_sdm, tor_unweighted_svm_occ)
plot(tor_unweighted_svm_occ_mask)
```

### Weighted spatial SVMs

For the weighted SVMs, I modified the code in the `svm.R` script of `maskRangeR` so that when `weight = TRUE`, the `e1071::svm()` function uses `class.weights = "inverse"`. According to the `e1071` documentation, this means the weights will be inversely proportional to the class distribution. We still need to discuss if this is the appropriate way to deal with weighting SVMs when there are more than 2 species.

```{r, message = FALSE}
# create SVM
weighted_svm_occ <- rangeSVM(variegatus[,2:3], tridactylus[,2:3], torquatus[,2:3], nrep = 100, weight = TRUE)
```

```{r}
# use SVM to create a raster of predicted regions
sloth_weighted_svm_occ <- rangeSVM_predict(svm = weighted_svm_occ, r = var_sdm)
plot(sloth_weighted_svm_occ)
points(variegatus[,2:3], pch = 20, cex = 0.75, col = "black")
points(tridactylus[,2:3], pch = 20, cex = 0.75, col = "red")
points(torquatus[,2:3], pch = 20, cex = 0.75, col = "blue")
```

#### *B. variegatus*

```{r}
# masked SDM predictions for variegatus
var_weighted_svm_occ <- sloth_weighted_svm_occ == 1
var_weighted_svm_occ[var_weighted_svm_occ == 0] <- NA
var_weighted_svm_occ_mask <- mask(var_sdm, var_weighted_svm_occ)
plot(var_weighted_svm_occ_mask)
```

#### *B. tridactylus*

```{r}
# masked SDM predictions for tridactylus
tri_weighted_svm_occ <- sloth_weighted_svm_occ == 2
tri_weighted_svm_occ[tri_weighted_svm_occ == 0] <- NA
tri_weighted_svm_occ_mask <- mask(tri_sdm, tri_weighted_svm_occ)
plot(tri_weighted_svm_occ_mask)
```

#### *B. torquatus*

```{r}
# masked SDM predictions for torquatus
tor_weighted_svm_occ <- sloth_weighted_svm_occ == 3
tor_weighted_svm_occ[tor_weighted_svm_occ == 0] <- NA
tor_weighted_svm_occ_mask <- mask(tor_sdm, tor_weighted_svm_occ)
plot(tor_weighted_svm_occ_mask)
```

## Hybrid SVMs

The hybrid SVMs are created using the habitat suitability predicted by the species distribution models in addition to spatial information, i.e. occurrences of the species.

### Unweighted hybrid SVMs

```{r, message = FALSE}
# create SVM
unweighted_svm_hyb <- rangeSVM(variegatus[,2:3], tridactylus[,2:3], torquatus[,2:3], sdm = raster::stack(var_sdm, tri_sdm, tor_sdm), nrep = 100)
```

```{r}
sloth_unweighted_svm_hyb <- rangeSVM_predict(svm = unweighted_svm_hyb, r = var_sdm, sdm = raster::stack(var_sdm, tri_sdm, tor_sdm))
plot(sloth_unweighted_svm_hyb)
points(variegatus[,2:3], pch = 20, cex = 0.75, col = "black")
points(tridactylus[,2:3], pch = 20, cex = 0.75, col = "red")
points(torquatus[,2:3], pch = 20, cex = 0.75, col = "blue")
```

#### *B. variegatus*

```{r}
# masked SDM predictions for variegatus
var_unweighted_svm_hyb <- sloth_unweighted_svm_hyb == 1
var_unweighted_svm_hyb[var_unweighted_svm_hyb == 0] <- NA
var_unweighted_svm_hyb_mask <- mask(var_sdm, var_unweighted_svm_hyb)
plot(var_unweighted_svm_hyb_mask)
```

#### *B. tridactylus*

```{r}
# masked SDM predictions for tridactylus
tri_unweighted_svm_hyb <- sloth_unweighted_svm_hyb == 2
tri_unweighted_svm_hyb[tri_unweighted_svm_hyb == 0] <- NA
tri_unweighted_svm_hyb_mask <- mask(tri_sdm, tri_unweighted_svm_hyb)
plot(tri_unweighted_svm_hyb_mask)
```

#### *B. torquatus*

```{r}
# masked SDM predictions for torquatus
tor_unweighted_svm_hyb <- sloth_unweighted_svm_hyb == 3
tor_unweighted_svm_hyb[tor_unweighted_svm_hyb == 0] <- NA
tor_unweighted_svm_hyb_mask <- mask(tor_sdm, tor_unweighted_svm_hyb)
plot(tor_unweighted_svm_hyb_mask)
```

### Weighted hybrid SVMs

```{r, message = FALSE}
# create SVM
weighted_svm_hyb <- rangeSVM(variegatus[,2:3], tridactylus[,2:3], torquatus[,2:3], sdm = raster::stack(var_sdm, tri_sdm, tor_sdm), nrep = 100, weight = TRUE)
```

```{r}
sloth_weighted_svm_hyb <- rangeSVM_predict(svm = weighted_svm_hyb, r = var_sdm, sdm = raster::stack(var_sdm, tri_sdm, tor_sdm))
plot(sloth_weighted_svm_hyb)
points(variegatus[,2:3], pch = 20, cex = 0.75, col = "black")
points(tridactylus[,2:3], pch = 20, cex = 0.75, col = "red")
points(torquatus[,2:3], pch = 20, cex = 0.75, col = "blue")
```

#### *B. variegatus*

```{r}
# masked SDM predictions for variegatus
var_weighted_svm_hyb <- sloth_weighted_svm_hyb == 1
var_weighted_svm_hyb[var_weighted_svm_hyb == 0] <- NA
var_weighted_svm_hyb_mask <- mask(var_sdm, var_weighted_svm_hyb)
plot(var_weighted_svm_hyb_mask)
```

#### *B. tridactylus*

```{r}
# masked SDM predictions for tridactylus
tri_weighted_svm_hyb <- sloth_weighted_svm_hyb == 2
tri_weighted_svm_hyb[tri_weighted_svm_hyb == 0] <- NA
tri_weighted_svm_hyb_mask <- mask(tri_sdm, tri_weighted_svm_hyb)
plot(tri_weighted_svm_hyb_mask)
```

#### *B. torquatus*

```{r}
# masked SDM predictions for torquatus
tor_weighted_svm_hyb <- sloth_weighted_svm_hyb == 3
tor_weighted_svm_hyb[tor_weighted_svm_hyb == 0] <- NA
tor_weighted_svm_hyb_mask <- mask(tor_sdm, tor_weighted_svm_hyb)
plot(tor_weighted_svm_hyb_mask)
```



